
\section{Concept of Differential Equations}

\begin{definition}{Differential Equation}{}
  A \emph{differential equation} is an equation that relates a function
  with one or more of its derivatives.
\end{definition}

\begin{definition}{Ordinary Differential Equation}{}
  An \emph{ordinary differential equation} involves a function
  of one independent variable and its derivatives.
\end{definition}

\begin{definition}{Partial Differential Equation}{}
  A \emph{partial differential equation} involves a function of
  multiple independent variables and its partial derivatives.
\end{definition}

\begin{definition}{Order of ODEs}{}
  The \emph{order of a ODE} is the highest derivative of the dependent variable
  with respect with the independent variable that appears in the equation.
\end{definition}

\begin{definition}{Linear ODEs}{}
  An \emph{$n$-th order linear ODE} is an ODE of the form
  \begin{equation}
    a_n(x)y^{(n)}(x) + \cdots + a_1(x)y^{\prime}(x) + a_0(x)y = g(x).
  \end{equation}
  If $a_i(x)$ are constant, then we say it is a \emph{constant-coefficient linear ODE}.
  If $g(x) \equiv 0$, then we say it is a \emph{homogeneous linear ODE}.
\end{definition}

\section{Solution of ODEs}

\begin{definition}{General Solution}{}
  The \emph{general solution} of an ODE is the form of the solution that includes all possible solutions.
\end{definition}

\begin{definition}{Particular Solution}{}
  A \emph{particular solution} of an ODE is a specific solution obtained from the general solution
  by assigning values to the arbitrary constants.
\end{definition}

\section{Forms of ODEs}

\begin{definition}{Explicit Form}{}
  In \emph{explicit form},
  the highest-order derivative is explicitly expressed in terms of the independent variable
  and the dependent variable (and possibly lower-order derivatives).
\end{definition}

\begin{definition}{Implicit Form}{}
  In \emph{implicit form},
  the equation relates the dependent variable, its derivatives, and the independent variable
  without explicitly solving for any derivative.
\end{definition}







